```{r setup, echo=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(kableExtra)
library(HelpersMG)
library(tidyr)
library(dplyr)
library(lubridate)
library(scales)
```
 <!-- This chunk is a library loader --> 

```{r load_data,cache=TRUE,echo=FALSE,message=FALSE,warning=FALSE}
wget ("https://covid.ourworldindata.org/data/owid-covid-data.csv", destfile="./raw_data/owid-covid-data.csv")
```
 <!-- This is for loading the owid dataset --> 

```{r,echo=FALSE,message=FALSE,warning=FALSE}
Diabetestibble<-read_csv("./raw_data/owid-covid-data.csv") %>% select(location, diabetes_prevalence,total_cases,date)  %>% write_csv("./output/diabetes_total_cases.csv")
```

 <!-- diabetes related to cases -->  

```{r,echo=FALSE,message=FALSE,warning=FALSE}
Diadeaths<-read_csv("./raw_data/owid-covid-data.csv") %>% select(location, diabetes_prevalence,total_deaths,date)  %>% write_csv("./output/diabetes_total_deaths.csv")
```

 <!-- diabetes related to deaths -->  
 
```{r,echo=FALSE,message=FALSE,warning=FALSE}
Cardiocases<-read_csv("./raw_data/owid-covid-data.csv") %>% select(location, cardiovasc_death_rate,total_cases,date)  %>% write_csv("./output/cardio_total_cases.csv")
```
 
 <!-- cardiovascular deaths to total cases -->  
 
```{r,echo=FALSE,message=FALSE,warning=FALSE}
Cardiodeaths<-read_csv("./raw_data/owid-covid-data.csv") %>% select(location, cardiovasc_death_rate,total_deaths,date)  %>% write_csv("./output/cardio_total_deaths.csv")
```
 
 
 <!-- Cardiovascular Deaths to total deaths --> 
 
##  Introduction

The aim of the following studies was to see if there was sufficient correlation between Respiratory Ailments and COVID-19 deaths and cases.
 


## Section 1: Deaths with Underlying Conditions
 
This section deals with death from COVID-19 where a underlying condition was also observed. The figured were derived from the CDC Dataset on underlying conditions. And the CDC dataset of Provisional Death Counts.
 
```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',fig.cap="__Figure 1__: Deaths with Underlying Conditions"}
deaths_underlying<-read_csv("./raw_data/conditions_deaths.csv") %>% select(State,Condition,`Number of COVID-19 Deaths`,`Age Group`) %>% filter(`Age Group` =="All ages" & State=="US",Condition !="COVID-19",Condition != "All other conditions and causes (residual)") %>% write_csv("./output/underlying_Causes_death.csv") %>% mutate(Respiratory = if_else(Condition %in% c("Adult respiratory distress syndrome", "Chronic lower respiratory diseases", "Influenza and pneumonia", "Other diseases of the respiratory system", "Respiratory arrest", "Respiratory failure"), TRUE, FALSE))

ggplot(deaths_underlying, aes(reorder(Condition, -`Number of COVID-19 Deaths`),fill = Respiratory)) + geom_bar(aes(weight = `Number of COVID-19 Deaths`)) + theme(axis.text.x = element_text(angle = 75, vjust = 1, hjust = 1))+labs(y="Deaths", x="Conditions")+ggsave("./figures/Underlying_conditions.png")
knitr::kable(anova(lm(`Number of COVID-19 Deaths`~ `Respiratory`, data = deaths_underlying)),
digits = 3,
format = "html",
caption = "Table 2: Provisional Death Anova",
table.attr = "style='width:70%;'") %>% kable_styling(position = "center")
```
 <!-- This is plot of the CDC data on underlying conditions and death --> 

This figure shows key factors causing the most deaths with COVID 19 are respiratory based. Influenza/ pneumonia take the top spot with Respiratory failure coming in second. Below is a table so you can also analyze the hard numbers.


```{r,echo=FALSE,message=FALSE,fig.align='center'}
knitr::kable(deaths_underlying %>% select(-c(State, `Age Group`)) %>% arrange(desc(`Number of COVID-19 Deaths`))) %>% kable_styling(full_width=FALSE)
```
 <!-- This chunk for the underlying deaths table --> 




```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center'}
provisional_death_counts<-read_csv("./raw_data/Provisional_Death_Counts.csv") %>% select(State,Indicator,`Start week`,`All COVID-19 Deaths (U07.1)`,`Deaths with Pneumonia and COVID-19 (J12.0-J18.9 and U07.1)`) %>% filter(Indicator == "Week-ending") %>% rename(`All COVID-19 Deaths`=`All COVID-19 Deaths (U07.1)`, `Deaths with Pneumonia and COVID-19`=`Deaths with Pneumonia and COVID-19 (J12.0-J18.9 and U07.1)`) %>% write_csv("./output/provionsal_death_counts.csv")

knitr::kable(provisional_death_counts %>% select(-c(State, Indicator)) %>% filter(`All COVID-19 Deaths` != 0) %>% mutate(`Percent of COVID Deaths involving Pneumonia` = signif(100.0 * `Deaths with Pneumonia and COVID-19` / `All COVID-19 Deaths`, 4))) %>% kable_styling(full_width=FALSE)


```
 <!-- This is code for the provisional death counts table --> 
 
This table shows a high correlation between the amount of COVID-19 deaths per week and that patient also having pneumonia. At times nearly half the patients who died from COVID-19 in a week also had pneumonia 


```{r}
knitr::kable(anova(lm(`All COVID-19 Deaths`~ `Deaths with Pneumonia and COVID-19`, data = read_csv("./output/provionsal_death_counts.csv"))),
			 digits = 3, 
			 format = "html", 
			 caption = "Table 2: Provisional Death Anova",
			 table.attr = "style='width:70%;'") %>% kable_styling(position = "center")
```
 <!-- Chunk for the ANOVA analysis --> 
 
 The ANOVA reveals a very high F value. You would expect this value to be around 1 but a large value means the null hypothesis is wrong.


## Section 2: Hospital Visits vs. Underlying Conditions  

In this section we will explore if there is a correlation between COVID-19 hospitalizations and the underlying conditions that accompanied them. The data was derived from the CDC Covid-Net datasets on medical conditions for aduts and for children 
```{r,echo=FALSE,message=FALSE}
medical_conditions<-read_csv("./raw_data/medical_conditions.csv", skip=2) %>% select(`MEDICAL CONDITION`,`ADULTS-COUNT`) %>% filter(`MEDICAL CONDITION`!="Pregnancy") %>% mutate(`ADULTS-COUNT` = as.integer(`ADULTS-COUNT`)) %>% write_csv("./output/hospitalization_causes.csv") %>% rename(`Adult Visits`="ADULTS-COUNT")
```
 <!-- This chunk is for comparing hospital adult with underlying causes --> 

```{r,echo=FALSE,message=FALSE,fig.align='center',fig.cap="__Figure 2__: Hospital Visits for Adults with accompanying conditions."}

ggplot(medical_conditions %>% filter(`MEDICAL CONDITION` != "No known condition" & `MEDICAL CONDITION` != "Other disease"), aes(reorder(`MEDICAL CONDITION`, -`Adult Visits`), `Adult Visits`)) + xlab("Medical Condition")  + geom_bar(stat="identity")+theme(axis.text.x = element_text(angle = 75, vjust = 1, hjust = 1))+ggsave("./figures/Adult_Hospital_Visits.png")
```



In this case we see in the COVID Center data that among hospitalizations in the US that hypertension was the largest underlying condition observed for adults. Respiratory afflictions like Asthma did not rank very highly. Below is a table of these results


```{r,echo=FALSE,message=FALSE,fig.align='center'}
knitr::kable(medical_conditions %>% filter(`MEDICAL CONDITION` != "No known condition" & `MEDICAL CONDITION` != "Other disease") %>% arrange(desc(`Adult Visits`))) %>% kable_styling(full_width=FALSE)

```
 <!-- This chunk is for the table for medical conditions --> 




```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.align='center',fig.cap="__Figure 3__: COVID-19 Children Hospitilaztions with accompanying conditions."}
kidhospital<-read_csv("./raw_data/medical_conditions.csv", skip=2) %>% select(`MEDICAL CONDITION`,`CHILDREN-COUNT`) %>% filter(`MEDICAL CONDITION`!="Pregnancy"&`MEDICAL CONDITION`!="No known condition"&`MEDICAL CONDITION`!="Other disease")%>% mutate(`CHILDREN-COUNT` = as.integer(`CHILDREN-COUNT`)) %>% rename(`Child Visits`="CHILDREN-COUNT") %>% write_csv("./output/kidhopsitalization.csv")
ggplot(kidhospital, aes(reorder(`MEDICAL CONDITION`, -`Child Visits`), y=`Child Visits`)) + geom_bar(stat="identity")+theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1))+ggsave("./figures/Child_Hospital_Visits.png")

```

 <!-- THis is for ailments versus Child hospitalization -->  
 
 The most likely underlying cause when it came to children was having no known other causes to go along with the COVID 19 diagnosis. 
 
```{r,echo=FALSE,message=FALSE,fig.align='center'}
knitr::kable(kidhospital %>% arrange(desc(`Child Visits`))) %>% kable_styling(full_width=FALSE)
```
 




 


